
Reproducing Variant Annotation Results
---------------------------------------

Below we compare the results of annotating variants via the [VariantAnnotation](http://www.bioconductor.org/packages/release/bioc/html/VariantAnnotation.html)  [BioConductor](http://www.bioconductor.org/) package.  We compare using data from 1,000 Genomes Phase 1 Variants:
* as parsed from a VCF file
* retrieved from the Google Genomics API

### VCF Data

First we read in the data from the VCF file:
```{r}
require(VariantAnnotation)
fl <- system.file("extdata", "chr22.vcf.gz", package="VariantAnnotation")
vcf <- readVcf(fl, "hg19")
vcf
```

The file `chr22.vcf.gz` within package VariantAnnotation holds data for 5 of the 1,092 individuals in 1,000 Genomes, starting at position 50300078 and ending at position 50999964.

`HG00096 HG00097 HG00099 HG00100 HG00101`

### Google Genomics Data

Important data differences to note:
 1 VCF data is 1-based but data from the GA4GH APIs is 0-based, for more information see (https://www.biostars.org/p/84686/)
 1 There are two variants in the Google Genomics copy of 1,000 Genomes phase 1 variants that are not in `chr22.vcf.gz`.  They are the only two variants within the genomic range tiwh `ALT == <DEL>`.

```{r}
require(GoogleGenomics)

# TODO Right now we're just getting a few variants.  Later update this to retrieve them all.
system.time({
granges <- getVariants(datasetId="10473108253681171589",
                       chromosome="22",
                       start=50300078,
                       end=50303000,         # TODO end=50999964
                       oneBasedCoord=TRUE,
                       converter=variantsToGRanges)
})
```

### Compare the Loaded Data
Ensure that the data retrieved by each matches:
```{r}
vcf <- vcf[1:length(granges)] # Truncate the VCF data

require(testthat)
expect_equal(start(granges), start(vcf))
expect_equal(end(granges), end(vcf))
expect_equal(as.character(granges$REF), as.character(ref(vcf)))
expect_equal(as.character(unlist(granges$ALT)), as.character(unlist(alt(vcf))))
expect_equal(granges$QUAL, qual(vcf))
expect_equal(granges$FILTER, filt(vcf))
```

### Compare the Annotations
Now locate the protein coding variants in each:
```{r}
library(TxDb.Hsapiens.UCSC.hg19.knownGene)
txdb <- TxDb.Hsapiens.UCSC.hg19.knownGene

vcf <- renameSeqlevels(vcf, c("22"="chr22"))
granges <- renameSeqlevels(granges, c("22"="chr22"))

rd <- rowData(vcf)
vcf_locations <- locateVariants(rd, txdb, CodingVariants())
vcf_locations

granges_locations <- locateVariants(granges, txdb, CodingVariants())
granges_locations

expect_equal(granges_locations, vcf_locations)
```

And predict the effect of the protein coding variants:
```{r}
library(BSgenome.Hsapiens.UCSC.hg19)
vcf_coding <- predictCoding(vcf, txdb, seqSource=Hsapiens)
vcf_coding

granges_coding <- predictCoding(rep(granges, elementLengths(granges$ALT)),
                                txdb,
                                seqSource=Hsapiens,
                                varAllele=unlist(granges$ALT, use.names=FALSE))

granges_coding

expect_equal(as.matrix(granges_coding$REFCODON), as.matrix(vcf_coding$REFCODON))
expect_equal(as.matrix(granges_coding$VARCODON), as.matrix(vcf_coding$VARCODON))
expect_equal(granges_coding$GENEID, vcf_coding$GENEID)
expect_equal(granges_coding$CONSEQUENCE, vcf_coding$CONSEQUENCE)
```

### Provenance
Package versions used:
```{r}
sessionInfo()
```

